Personality and political beliefs across the lifespan

### 2019-12-11 scripts/manuscript.Rmd
cleaning up folders

---
title             : "Personality and political beliefs across the lifespan"
shorttitle        : "PERSONALITY AND POLITICAL BELIEFS"
author: 
  - name          : "Sarah Dimakis"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "University of Oregon, Eugene, OR"
    email         : "sdimakis@uoregon.edu"
  - name          : "Meghan Siritzky"
    affiliation   : "1"
  - name          : "Jamie Yellowtail"
    affiliation   : "1"
affiliation:
  - id            : "1"
    institution   : "University of Oregon"
authornote: |
  This project was completed as part of the EDLD Introduction to Data Science class at the University of Oregon.
abstract: |
  Since the development of the Five Factor Model as a framework for measuring and classifying personality, psychology research has been interested in using this framework to examine the relationship between personality and cognitions, mindsets, and behaviors. Prior research has attempted to predict political beliefs and affiliations from personality traits, finding that the personality trait of openness to experience is positively correlated with liberalness, while the trait of conscientiousness is positively correlated with conservativeness. The current study attempts to replicate these findings in a U.S. Amazon Mechanical Turk population using the Ten Item Personality Inventory to measure the Big Five personality traits (openness to experiences, conscientiousness, extraversion, agreeableness, and emotional stability) and a 12-item Social and Economic Conservatism Scale to measure the social and economic dimensions of conservatism. Consistent with past research, the current study found that openness to experience was a significant negative predictor of both economic and social conservatism. Contrary to past findings, however, we found that agreeableness and emotional stability were both significant positive predictors of social conservatism, and conscientiousness was not a significant predictor of either social or economic conservatism. Neither social nor economic extraversion were strongly correlated with any of the Big Five personality traits, but they were strongly positively correlated with each other.

@@

```{r setup, include = FALSE}
library(papaja)
library(rio)
library(here)
library(tidyverse)
library(psych)
library(knitr)
library(MBESS)
library(viridis)
library(corrplot)

@@

Consistent with prior research on personality and political beliefs,  participants in this study  that were higher in economic conservatism produced significantly lower scores on openness to experience. This relationship has been described in prior research as a result of the ideas and beliefs surrounding economic conservatism, such as rule following, self-regulation, and order, contradicting the ideas and beliefs surrounding openness to experience, such as creativity and diversity of experience [@carney2008secret; @ekehammar2004matters].

>>>>

---
title             : "Personality and political beliefs across the lifespan"
shorttitle        : "PERSONALITY AND POLITICAL BELIEFS"
author: 
  - name          : "Sarah Dimakis"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "University of Oregon, Eugene, OR"
    email         : "sdimakis@uoregon.edu"
  - name          : "Meghan Siritzky"
    affiliation   : "1"
  - name          : "Jamie Yellowtail"
    affiliation   : "1"
affiliation:
  - id            : "1"
    institution   : "University of Oregon"
authornote: |
  This project was completed as part of the EDLD Introduction to Data Science class at the University of Oregon.
abstract: |
  Since the development of the Five Factor Model as a framework for measuring and classifying personality, psychology research has been interested in using this framework to examine the relationship between personality and cognitions, mindsets, and behaviors. Prior research has attempted to predict political beliefs and affiliations from personality traits, finding that the personality trait of openness to experience is positively correlated with liberalness, while the trait of conscientiousness is positively correlated with conservativeness. The current study attempts to replicate these findings in a U.S. Amazon Mechanical Turk population using the Ten Item Personality Inventory to measure the Big Five personality traits (openness to experiences, conscientiousness, extraversion, agreeableness, and emotional stability) and a 12-item Social and Economic Conservatism Scale to measure the social and economic dimensions of conservatism. Consistent with past research, the current study found that openness to experience was a significant negative predictor of both economic and social conservatism. Contrary to past findings, however, we found that agreeableness and emotional stability were both significant positive predictors of social conservatism, and conscientiousness was not a significant predictor of either social or economic conservatism. Neither social nor economic conservatism were strongly correlated with any of the Big Five personality traits, but they were strongly positively correlated with each other.

@@

```{r setup, include = FALSE}
#install.packages(c("corrplot", "MBESS", "psych"))
library(papaja)
library(rio)
library(here)
library(tidyverse)
library(psych)
library(knitr)
library(MBESS)
library(viridis)
library(corrplot)

@@

Consistent with prior research on personality and political beliefs,  participants in this study  that were higher in economic conservatism produced significantly lower scores on openness to experience. This relation has been described in prior research as a result of the ideas and beliefs surrounding economic conservatism, such as rule following, self-regulation, and order, contradicting the ideas and beliefs surrounding openness to experience, such as creativity and diversity of experience [@carney2008secret; @ekehammar2004matters].

### 2019-12-11 scripts/manuscript.tex
cleaning up folders

\abstract{
Here is where we will write our abstract.

@@

We used R (Version 3.6.1; R Core Team, 2019) and the R-packages
\emph{\}MBESS} {[}@\}R-MBESS{]}, \emph{dplyr} (Version 0.8.3; Wickham,
François, Henry, \& Müller, 2019), \emph{forcats} (Version 0.4.0;
Wickham, 2019a), \emph{ggplot2} (Version 3.2.1; Wickham, 2016),
\emph{here} (Version 0.1; Müller, 2017), \emph{janitor} (Version 1.2.0;
Firke, 2019), \emph{knitr} (Version 1.25; Xie, 2015), \emph{papaja}
(Version 0.1.0.9842; Aust \& Barth, 2018), \emph{psych} (Version 1.8.12;
Revelle, 2018), \emph{purrr} (Version 0.3.3; Henry \& Wickham, 2019),
\emph{readr} (Version 1.3.1; Wickham, Hester, \& Francois, 2018),
\emph{rio} (Version 0.5.16; C.-h. Chan, Chan, Leeper, \& Becker, 2018),
\emph{stringr} (Version 1.4.0; Wickham, 2019b), \emph{tibble} (Version
2.1.3; Müller \& Wickham, 2019), \emph{tidyr} (Version 1.0.0; Wickham \&
Henry, 2019), \emph{tidyverse} (Version 1.2.1; Wickham, 2017),
\emph{viridis} (Version 0.5.1; Garnier, 2018a, 2018b), and
\emph{viridisLite} (Version 0.3.0; Garnier, 2018b) for all our analyses.

@@

@@

@@

@@

\hypertarget{ref-R-forcats}{}
Wickham, H. (2019a). \emph{Forcats: Tools for working with categorical
variables (factors)}. Retrieved from
\url{https://CRAN.R-project.org/package=forcats}

\hypertarget{ref-R-stringr}{}
Wickham, H. (2019b). \emph{Stringr: Simple, consistent wrappers for
common string operations}. Retrieved from
\url{https://CRAN.R-project.org/package=stringr}

>>>>

\abstract{
Since the development of the Five Factor Model as a framework for
measuring and classifying personality, psychology research has been
interested in using this framework to examine the relationship between
personality and cognitions, mindsets, and behaviors. Prior research has
attempted to predict political beliefs and affiliations from personality
traits, finding that the personality trait of openness to experience is
positively correlated with liberalness, while the trait of
conscientiousness is positively correlated with conservativeness. The
current study attempts to replicate these findings in a U.S. Amazon
Mechanical Turk population using the Ten Item Personality Inventory to
measure the Big Five personality traits (openness to experiences,
conscientiousness, extraversion, agreeableness, and emotional stability)
and a 12-item Social and Economic Conservatism Scale to measure the
social and economic dimensions of conservatism. Consistent with past
research, the current study found that openness to experience was a
significant negative predictor of both economic and social conservatism.
Contrary to past findings, however, we found that agreeableness and
emotional stability were both significant positive predictors of social
conservatism, and conscientiousness was not a significant predictor of
either social or economic conservatism. Neither social nor economic
conservatism were strongly correlated with any of the Big Five
personality traits, but they were strongly positively correlated with
each other.

@@

We used R (Version 3.6.1; R Core Team, 2019) and the R-packages
\emph{\}forcats} {[}@\}R-forcats{]}, \emph{\}MBESS} {[}@\}R-MBESS{]},
\emph{corrplot2017} (Wei \& Simko, 2017), \emph{dplyr} (Version 0.8.3;
Wickham, François, Henry, \& Müller, 2019), \emph{GGally} (Schloerke et
al., 2018), \emph{ggplot2} (Version 3.2.1; Wickham, 2016), \emph{here}
(Version 0.1; Müller, 2017), \emph{janitor} (Version 1.2.0; Firke,
2019), \emph{knitr} (Version 1.25; Xie, 2015), \emph{papaja} (Version
0.1.0.9842; Aust \& Barth, 2018), \emph{psych} (Version 1.8.12; Revelle,
2018), \emph{purrr} (Version 0.3.3; Henry \& Wickham, 2019),
\emph{readr} (Version 1.3.1; Wickham, Hester, \& Francois, 2018),
\emph{rio} (Version 0.5.16; C.-h. Chan, Chan, Leeper, \& Becker, 2018),
\emph{stringr} (Version 1.4.0; Wickham, 2019), \emph{tibble} (Version
2.1.3; Müller \& Wickham, 2019), \emph{tidyr} (Version 1.0.0; Wickham \&
Henry, 2019), \emph{tidyverse} (Version 1.2.1; Wickham, 2017),
\emph{viridis} (Version 0.5.1; Garnier, 2018a, 2018b), and
\emph{viridisLite} (Version 0.3.0; Garnier, 2018b) for all our analyses.

@@

Consistent with prior research on personality and political beliefs,
participants in this study that were higher in economic conservatism
produced significantly lower scores on openness to experience. This
relation has been described in prior research as a result of the ideas
and beliefs surrounding economic conservatism, such as rule following,
self-regulation, and order, contradicting the ideas and beliefs
surrounding openness to experience, such as creativity and diversity of
experience (Carney et al., 2008; Ekehammar et al., 2004).

However, inconsistent with prior research and our hypotheses, the
negative relationship between social conservatism and openness to
experience scores, along with the positive relationships between
conscientiousness and both social and economic conservatism were not
significant. With this subject so heavily studied, the lack of
significant results indicates that a replication of the study that
includes a power analysis and possibly, an increase in sample size may
give us a clearer understanding of the strength and significance of
these relationships.

Additional exploratory analysis revealed a significant positive
relationship between emotional stability and social conservatism, as
well as a positive relationship between agreeableness and social
conservatism. Prior research indicates that emotional stability is not a
significant predictor of conservatism, and agreeableness is usually
negatively correlated with conservatism (Sibley, Osborne, \& Duckitt,
2012). The departure of these results from previous findings may
indicate that there is an increase in positive personality traits
associated with conservatism, but the controlling, firm, inhibited
characteristics heavily associated with conservatism are not compatible
with the change in direction between agreeableness and conservatism.

While the results of this study did not support all of our hypotheses on
personality traits and conservatism, the limitations of the study
surrounding the self-reporting methods along with the potential low
power of the study may help explain the departure of our results from
previous findings. Future research on this subject would be more
informed through the use of experimental methods and observations of
participants to determine behavior that reveals personality traits and
level of social and economic conservatism. Increasing understanding of
the relationship between personality traits and left-right differences
could help implement actions that would help diffuse the increasing
political polarization currently taking place in the United States.

@@

\hypertarget{ref-R-GGally}{}
Schloerke, B., Crowley, J., Cook, D., Briatte, F., Marbach, M., Thoen,
E., \ldots{} Larmarange, J. (2018). \emph{GGally: Extension to
'ggplot2'}. Retrieved from
\url{https://CRAN.R-project.org/package=GGally}

\hypertarget{ref-sibley2012personality}{}
Sibley, C. G., Osborne, D., \& Duckitt, J. (2012). Personality and
political orientation: Meta-analysis and test of a threat-constraint
model. \emph{Journal of Research in Personality}, \emph{46}(6),
664--677.

@@

\hypertarget{ref-R-corrplot2017}{}
Wei, T., \& Simko, V. (2017). \emph{R package ``corrplot'':
Visualization of a correlation matrix}. Retrieved from
\url{https://github.com/taiyun/corrplot}

@@

\hypertarget{ref-R-stringr}{}
Wickham, H. (2019). \emph{Stringr: Simple, consistent wrappers for
common string operations}. Retrieved from
\url{https://CRAN.R-project.org/package=stringr}
